Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex 

### 2019-01-30 Manuscript.Rmd
Reshaping the paper with current data

With these lists, we identified areas that were significantly more associated to either literature. We limited the search space to the parcels that were contained within the medial portion of the DMN parcellation defined by Yeo et al. [-@Yeo2011], as it broadly encompassed our a priori domain of interest. We permuted the label of every study (i.e. DMN or valuation) 5000 times while preseving the total number of studies observed in each domain, and on each iteration stored the maximum resulting statistic from chi-squared tests of proportions that compared the number of studies that contained each area. This gave us a null distribution of 5000 chi-squared values. We then used the 95th percentile of this distribution as a significance threshold to evaluate the chi-squared values produced by comparing the actual rates of reporting of each area between literatures. The resulting ROIs contained two spatially discontiguous regions (Figure 1), which we label PCC and mPFC for further analyses.

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,081 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and took the exponential of these values so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately (1200 TRs each); 2) the concatenated timeseries from each pair of daily sessions (2400 TRs total); 3) each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

Communities (i.e. clusters) were identified by means of spectral partitioning [@Fiedler1975; @Belkin2003]. First, each network was represented as an $n$ x $n$ matrix of previously adjusted correlation values (where $n$ = number of vertices within the ROIs, 5,319). This matrix was then transformed into its normalized Laplacian form

@@

BOLD signal recording through fMRI in mPFC in general, and vmPFC in particular, suffers from technical issues that promote signal dropout [@Logothetis]. It was thus important to quantify the potential influence of noise in our findings. We examined this through two complementary steps. First, we computed the temporal signal-to-noise ratio (tSNR) of all the vertices in the ROIs from HCP minimally preprocessed data (i.e. prior to the complementary denoising described in the "fMRI Data" section). We compared the mean tSNR for each region across subjects through a permutation analysis (all permutation analyses involved 5000 iterations). In addition, we compared the mean tSNR between mPFC and PCC through a permutation analysis for each separate subject, and report the percentage of subjects who had a significantly lower mean tSNR in mPFC after FDR correction for multiple testing [@Benjamini2001]. Next, we estimated which vertices within mPFC had a stronger relationship to the overall dynamics of DMN. To do this, we computed the mean time series of each subject's estimated DMN map, which was then predicted by the time series each mPFC vertex in a mass univariate analysis. The resulting $R^2$ values from each model were extracted to determine the total variance of the mean DMN time series explained by each individual vertex. If high noise in more ventral mPFC areas is affecting our results, we should see a gradient of $R^2$ values that decreases ventrally across subjects. We then spatially correlated (Pearson) the $R^2$ and tSNR values to determine whether low tSNR negatively impacted the individual contribution among vertices.

@@

# Permutation analysis
# Note: this doesn't care whether a region was uni or bilaterally activated. Maybe do that.
# Load the coordinates as studies in a list, because for the contrast the studies are permuted
# Use table(unlist(list)) to get the counts in that case
# setwd('./DEC')
# templist <- dir()
# studyList_sv <- lapply(templist, readr::read_csv, col_names = F)
# 
# setwd('../DMN/')
# templist <- dir()
# studyList_dmn <- lapply(templist, readr::read_csv, col_names = F)
# 
# # For each permutation, reassign the study labels, maintaining the original number of studies per condition
# allStudies <- c(studyList_sv, studyList_dmn)
# nPerms <- 1000 # stored statistics are computed @ 5k, but it's computationally demanding to do that.
# nullDist <- numeric()
# 
# for (i in seq(nPerms)) {
# 
#   if(i %in% c(nPerms*0.25, nPerms*0.5, nPerms*0.75, nPerms*0.95)) {print(paste("Completed", i/nPerms, "percent of the permutation..."))}
# 
#   # Sample all the studies and divide into 2 groups
#   randPerm <- sample(allStudies)
#   tmpSV <- randPerm[seq(length(studyList_sv))]
#   tmpDMN <- randPerm[-seq(length(studyList_sv))]
# 
#   # Count the number of times each region was reported in the literature
#   # Bilateral activations still count as 1, so we might want to reconsider that
#   tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                       SV = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpSV))}),
#                       DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpDMN))}))
# 
#   # Perform the prop test for each pair and get the max chi-squared stat
#   nullDist[i] <- max(
#                   apply(tmpDF[,2:3], 1, function(data) {
#                     prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
#                     }
#                   ), na.rm = T)
# 
# }
# 
# # Now do the same for the actual data, to see how it compares to the null dist
# tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                     SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
#                     DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
#                     propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
#                     propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))
# 
# # Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
# tmpDF$Chi_squared <- apply(tmpDF[,2:3], 1, function(data) {
#   round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
#   })
# tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution
# 
# # ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
# significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist,0.95))

@@

We performed a coordinate-based meta-analysis to identify unique standard regions across DMN and SV literatures (see Methods). Volumetric coordinates from a set of DMN and SV studies were projected onto a cortical surface, and mapped to a multimodal cortical parcellation [@Glasser2016] to produce a list of brain areas reported per study. The areas considered were limited to the medial portion of the DMN, as defined by Yeo et al.[-@Yeo2011]. Unique areas were identified by permuting the affiliation of each study (i.e. DMN or SV) to create a null distribution of chi-squared statistics, which was used to identify regions more often reported in either literature. Importantly, we chose to examine these areas bilaterally, even if their significance was present in only one hemisphere.
Figure 1 shows the resulting ROIs from these analyses. Area 7m was the only one to show unique affiliation, which was significant for DMN ($\chi^2$ = 9.12, *p* < 0.001).

@@

*NOTE: It might be better to just avoid cluster comparisons at this pre-threshold level*
With these 16 ROIs (15 overlapping, 1 DMN-specific), we estimated each individual's DMN coverage using the whole time series (4800 total TRs from four scanning sessions over two days). Each subject's network (n = `r nSubj`) was constructed by performing pairwise correlations among the time series of all the surface vertices contained within the ROIs (`r nrow(Summaries[[1]])` total vertices). After correcting for multiple testing, we performed spectral partitioning and modularity on these networks to identify cohesive functional communities denoting DMN and non-DMN areas. On average, networks had high densities (i.e. highly connected) regardless of the amount of data to produce them (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`). This degree of density is important, as these methods are known to face challenges when computed on sparse networks [@Fortunato2016].

Figure 2 shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning. This method dissects a network by the positive and negative values derived from the Fiedler vector (the eigenvector related to the second-to-lowest eigenvalue of the network's Laplacian matrix, see Methods), which we used to binarize the ROIs. Since this is a data-driven approach, there is no a priori labeling for these communities. We thus defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (7m), and ensured that the binary labeling of the vertices was consistent across subjects. In general, the resulting pattern shows large DMN coverage in PCC, with non-DMN vertices populating its anterior segment. mPFC, on the other hand, is mostly covered by DMN vertices in its ventral-anterior and dorsal-anterior areas, with a persistent non-DMN pocket between them. This non-DMN section extends posteriorly onto pregenual cinglate cortex. Note that vmPFC does not encompass area a25, which is often related to negative emotional valuation [@Joyce2018]. Interstingly, this topographical combination of dorsal and ventral mPFC, with extensive PCC coverage, matches previous meta-analytic patterns more closely attributed to social cognition than SV [@Hiser2018].

>>>>

With these lists, we identified areas that were significantly more associated to either literature. We limited the search space to the parcels that were contained within the medial portion of the DMN parcellation defined by Yeo et al. [-@Yeo2011], as it broadly encompassed our a priori domain of interest (25 parcels across hemispheres). We permuted the label of every study (i.e. DMN or valuation) 5000 times while preseving the total number of studies observed in each domain, and on each iteration stored the maximum resulting statistic from chi-squared tests of proportions that compared the number of studies that contained each area. This gave us a null distribution of 5000 chi-squared values. We then used the 95th percentile of this distribution as a significance threshold to evaluate the chi-squared values produced by comparing the actual rates of reporting of each area between literatures. The resulting ROIs contained two spatially discontiguous regions (Figure 1), which we label PCC and mPFC for further analyses.

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 4,801 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and took the exponential of these values so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately (1200 TRs each); 2) the concatenated timeseries from each pair of daily sessions (2400 TRs total); 3) each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

Communities (i.e. clusters) were identified by means of spectral partitioning [@Fiedler1975; @Belkin2003]. First, each network was represented as an $n$ x $n$ matrix of previously adjusted correlation values (where $n$ = number of vertices within the ROIs, 4,801). This matrix was then transformed into its normalized Laplacian form

@@

BOLD signal recording through fMRI in mPFC in general, and vmPFC in particular, suffers from technical issues that promote signal dropout [@Logothetis]. Even though we made attempts to remove clear sources of noise through data preprocessing and stability analyses, it was still important to quantify the remaining potential influence of noise in our findings. We examined this through two complementary steps. First, we computed the temporal signal-to-noise ratio (tSNR) of all the vertices in the ROIs from HCP minimally preprocessed data (i.e. prior to the complementary denoising described in the "fMRI Data" section). We compared the mean tSNR for each region across subjects through a permutation analysis (all permutation analyses involved 5000 iterations). In addition, we compared the mean tSNR between mPFC and PCC through a permutation analysis for each separate subject, and report the percentage of subjects who had a significantly lower mean tSNR in mPFC after FDR correction for multiple testing [@Benjamini2001]. Next, we estimated which vertices within mPFC had a stronger relationship to the overall dynamics of DMN. To do this, we computed the mean time series of each subject's estimated DMN map, which was then predicted by the time series each mPFC vertex in a mass univariate analysis. The resulting $R^2$ values from each model were extracted to determine the total variance of the mean DMN time series explained by each individual vertex. If high noise in more ventral mPFC areas is affecting our results, we should see a gradient of $R^2$ values that decreases ventrally across subjects. We then spatially correlated (Pearson) the $R^2$ and tSNR values to determine whether low tSNR negatively impacted the individual contribution among vertices.

@@

# # Permutation analysis
# # Note: this doesn't care whether a region was uni or bilaterally activated. Maybe do that.
# # Load the coordinates as studies in a list, because for the contrast the studies are permuted
# # Use table(unlist(list)) to get the counts in that case
# setwd('./DEC')
# templist <- dir()
# studyList_sv <- lapply(templist, readr::read_csv, col_names = F)
# 
# setwd('../DMN/')
# templist <- dir()
# studyList_dmn <- lapply(templist, readr::read_csv, col_names = F)
# 
# # For each permutation, reassign the study labels, maintaining the original number of studies per condition
# allStudies <- c(studyList_sv, studyList_dmn)
# nPerms <- 1000 # stored statistics are computed @ 5k, but it's computationally demanding to do that.
# nullDist <- numeric()
# 
# for (i in seq(nPerms)) {
# 
#   if(i %in% c(nPerms*0.25, nPerms*0.5, nPerms*0.75, nPerms*0.95)) {print(paste("Completed", i/nPerms, "percent of the permutation..."))}
# 
#   # Sample all the studies and divide into 2 groups
#   randPerm <- sample(allStudies)
#   tmpSV <- randPerm[seq(length(studyList_sv))]
#   tmpDMN <- randPerm[-seq(length(studyList_sv))]
# 
#   # Count the number of times each region was reported in the literature
#   # Bilateral activations still count as 1, so we might want to reconsider that
#   tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                       SV = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpSV))}),
#                       DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpDMN))}))
# 
#   # Perform the prop test for each pair and get the max chi-squared stat
#   nullDist[i] <- max(
#                   apply(tmpDF[,2:3], 1, function(data) {
#                     prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
#                     }
#                   ), na.rm = T)
# 
# }
# 
# # Now do the same for the actual data, to see how it compares to the null dist
# tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                     SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
#                     DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
#                     propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
#                     propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))
# 
# # Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
# tmpDF$Chi_squared <- apply(tmpDF[,2:3], 1, function(data) {
#   round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
#   })
# tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution
# 
# # ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
# significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist,0.95))

# results!
# s32 (prportion sv = 0.6; proportion DMN = 0.2, Chi = 11.95, p < 0.01)
# 7m (prportion sv = 0.07; proportion DMN = 0.4, Chi = 8.45, p < 0.05)
# null chi <- 8.4

@@

We performed a coordinate-based meta-analysis to identify unique standard regions across DMN and SV literatures (see Methods). Volumetric coordinates from a set of DMN and SV studies were projected onto a cortical surface, and mapped to a multimodal cortical parcellation [@Glasser2016] to produce a list of brain areas reported per study. The 25 areas considered were limited to the medial portion of the DMN, as defined by Yeo et al.[-@Yeo2011]. Unique areas were identified by permuting the affiliation of each study (i.e. DMN or SV) to create a null distribution of chi-squared statistics, which was used to identify regions more often reported in either literature. Importantly, we chose to examine these areas bilaterally, even if their significance was present in only one hemisphere.

Figure 1 shows the resulting ROIs from these analyses. The 95th percentile of the permuted Chi-squared distribution was 8.4. Based on this threshold, area 7m was the only one to show unique affiliation to DMN (observed $\chi^2$ = 8.45, *p* < 0.05), while SV was uniquely represented by area s32 (observed $\chi^2$ = 11.95, *p* < 0.01).

@@

*NOTE: It might be better to just avoid cluster comparisons at this pre-threshold level*
With these 25 ROIs (23 overlapping, 1 DMN-specific, and 1 SV-specific), we estimated each individual's DMN coverage using the whole time series (4800 total TRs from four scanning sessions over two days). Each subject's network (n = `r nSubj`) was constructed by performing pairwise correlations among the time series of all the surface vertices contained within the ROIs (`r nVertices` total vertices). After correcting for multiple testing, we performed spectral partitioning and modularity on these networks to identify cohesive functional communities denoting DMN and non-DMN areas. On average, networks had high densities (i.e. highly connected) regardless of the amount of data to produce them (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`). This degree of density is important, as these methods are known to face challenges when computed on sparse networks [@Fortunato2016].

Figure 2 shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning. This method dissects a network by the positive and negative values derived from the Fiedler vector (the eigenvector related to the second-to-lowest eigenvalue of the network's Laplacian matrix, see Methods), which we used to binarize the ROIs. Since this is a data-driven approach, there is no a priori labeling for these communities. We thus defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (7m), and ensured that the binary labeling of the vertices was consistent across subjects. In general, the resulting pattern shows large DMN coverage in PCC, with non-DMN vertices populating its anterior segment. mPFC, on the other hand, is mostly covered by DMN vertices in its ventral-anterior and dorsal-anterior areas, with a persistent non-DMN pocket between them. This non-DMN section extends posteriorly onto pregenual cinglate cortex. Interstingly, this topographical combination of dorsal and ventral mPFC, with extensive PCC coverage, matches previous meta-analytic patterns more closely attributed to social cognition than SV [@Hiser2018].
